---
title: "Simple Linear Regression"
author: "Esse Julien Atto"
output: 
  html_document:
    code_folding: hide
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,message=FALSE,warning=F}
#loading libraries to use
library(dplyr)
library(datasets) #for cars data
library(ggplot2)
```

Simple linear regression is a statistical method for finding a linear relationship between an explanatory variable $X$ and a variable to be explained $y$:
\[y= \theta_0 + \theta_1 X.\]

# I. Data preparation and visualisation
```{r}
#Loading car data

data(cars)

```

The data **cars** in R package **datasets** give the speed of cars and the distances taken to stop, and  were recorded in the 1920s.

## 1. Visualization of ten rows chosen randomly among the 50 observations
```{r}
set.seed(1234)
cars[sample(1:50,10),]
```


## 2. Graph
```{r}
#% Plot Data
plt<-cars%>%
  ggplot(aes(x = speed, y = dist)) +
  geom_point(colour="red") +
  xlab("Speed (mph)") +
  ylab("Stopping distance (feet)") +
  ggtitle("Speed and Stopping Distances of Cars")


plt

```

# II. Linear regresson using Gradient Descent
## 1. Computation of the parameters
```{r}
#X and y
X<-cars$speed
y=cars$dist

# initializing fitting parameters
theta = cbind(numeric(2)) 

# choosing number of iterations
nb_iter = 10000

#loading  some functions... 
source("functions.R")

##############################################################
  #Ploting cost as function of number of iterations for some values of the learning rate.
  # and choose theta related to the better learning rate. 
  #Here,it seems that good values of alpha are between 0.0075 and 0.007.
  ALPHA<-c(0.0075,0.0074,0.0073,0.0072)
  COL<-c("red","blue","green","yellow")
  out<-list()

  for(i in 1:length(ALPHA)){
    #% Initializing  theta and running Gradient Descent 
    theta = matrix(0,2)
    Grad_Desc<- Grad_Desc_fct(X, y, theta, alpha = ALPHA[i], nb_iter)
    theta<-Grad_Desc$theta
    Cost_by_iter<-Grad_Desc$Cost_by_iter
    out[[i]]<-theta
    #% Plot the convergence graph for different values of alpha
    curr_i=i
    if(curr_i==1){
      plot(1:nrow(Cost_by_iter), Cost_by_iter, type="l",lwd=2,xlab='Number of iterations',ylab='Cost',col=COL[1],
           main="Convergence of gradient descent for a given learning rate",xlim = c(0,400))
    }else{
      lines(1:nrow(Cost_by_iter), Cost_by_iter, type="l",lwd=2,col=COL[i])
    }
  }
  legend("topright", legend = paste("alpha=",ALPHA), 
         lwd = 2, cex = 1.2, col = COL, lty = rep(1,4))
```


```{r,echo=FALSE}
names(out)<-paste(ALPHA)

```

If we choose $\alpha=0.0075$, then we have:
```{r,echo=FALSE}
theta<-out$`0.0075`
```

\[\theta=\begin{pmatrix} `r theta[1]`\\`r theta[2]` \end{pmatrix}\]

## 2. Plotting the linear fit
```{r}
#Plot the linear fit

plt+geom_abline(intercept = theta[1], slope = theta[2], color="blue", 
               size=1.5,)
```

## 3. Predictions
### a. Predict stoping distance for a speed of 21 mph.
```{r}
dist_for_21 = c(c(1, 21)%*%theta)
```

For speed = 21 mph (`r round(1.609344*21,2)` km/h), we predict a stopping distance of  **`r round( dist_for_21,2)`** feet (**`r round(dist_for_21*0.3048,2)`** m).

### b, Predict stoping distance for a  speed of 30 mph.
```{r}
dist_for_30 =c( c(1, 30)%*%theta)

```

For speed = 30 mph (`r round(1.609344*30,2)` km/h), we predict a stopping distance of  **`r round(dist_for_30,2)`** feet (**`r round(dist_for_30*0.3048,2)`** m).

# III. Simple Linear Regression using lm() function of R (Normal equation)

\[\theta=(X^TX)^{-1}X^Ty.\]

```{r}
model_lm<-lm(dist~speed,data = cars)

model_lm
```

We can see that the values of the fitted parameters are $\hat{\theta_0}=`r round(model_lm[['coefficients']][1],3)`$ and $\hat{\theta_1}=`r round(model_lm[['coefficients']][2],3)`$.

## 1. Summary of the model
```{r}
summary(model_lm)
```


## 2. Predictions
### a. Predict stoping distance for a speed of 21 mph.
```{r}
dist_for_21_ =predict.lm(model_lm,data.frame(speed=21))
```

For speed = 21 mph (`r round(1.609344*21,2)` km/h), we predict a stopping distance of  **`r round(dist_for_21_,2)`** feet (**`r round(dist_for_21_*0.3048,2)`** m).

### a. Predict stoping distance for a speed of 30 mph.
```{r}
dist_for_30_ =predict.lm(model_lm,data.frame(speed=30))

```

For speed = 30 mph (`r round(1.609344*30,2)` km/h), we predict a stopping distance of  **`r round(dist_for_30_,2)`** feet (**`r round(dist_for_30_*0.3048,2)`** m).
